---
title: Count and cluster colors for route {{ from }} to {{ to }} 
output: github_document
---

```{r setup}
library(tidyverse)
library(colorspace)
library(routes)
library(here)
set.seed(21978) # both subsampling and clustering are random, 
                # set a seed to get reproducible image
```


## Read in Streetview image manfest

```{r load-manifest}
route_images <- read_rds(here("{{{ shortname }}}", "images", "manifest.rds"))
```

## Get sample of pixels from images

Sample 50 pixels from each image:
```{r sample-pixels}
pixel_sample_rgb <- route_images %>% 
  pull(image) %>% 
  map_dfr(sample_pixels, sample_size = 50) %>% 
  with(RGB(R, G, B)) 
```

## Create cluster centers
```{r cluster-params}
cluster_space <- "LUV"
n_clusters <- 50
```

Clustering done in `r cluster_space` color space, with `r n_clusters` clusters:
```{r cluster-colors, fig.width = 10, fig.height = 3}
pixel_sample_space <- pixel_sample_rgb %>% 
  as(cluster_space)

clusters <- kmeans(pixel_sample_space@coords, centers = n_clusters)
centers <- match.fun(cluster_space)(clusters$center)

centers %>% hex() %>% pal()
```

## Count pixels in each cluster

```{r count-colors-sample} 
color_freq_sample <- count_colours(pixel_sample_rgb, centers,
  colorspace = cluster_space)
```

Histogram of counts based on sample:
```{r hist-colors-sample, fig.width = 4, fig.height = 7}
color_freq_sample %>% 
  ggplot(aes(reorder(hex, freq), freq, fill = hex)) +
    geom_col() +
  scale_fill_identity() +
  coord_flip() +
  theme_classic()
```

## Count pixels in each image

```{r count-colors-all}
route_images <- route_images %>% 
  mutate(
    pixels = map(image, read_pixels) %>% map(RGB),
    color_count = map(pixels, count_colours, 
      centers = centers, colorspace = "LUV")) %>% 
  unnest(color_count) %>% 
  mutate(H = as(LUV(L, U, V), "polarLUV")@coords[, "H"])
```

## Initial Plot

```{r initial-plot, fig.width = 10, fig.height = 3}
route_images %>% 
  ggplot(aes(order, freq)) +
    geom_area(aes(fill = hex)) +
  scale_fill_identity() +
  scale_color_identity() +
  equal_margins()
```

```{r order-param, fig.width = 10, fig.height = 3}
orderby <- quo(V)
```

You might try re-ordering the colors by one of the color dimensions, e.g. `r rlang::quo_text(orderby)`, other options are `H`, `U` or `L`:
```{r ordered-plot, fig.width = 10, fig.height = 3}
route_images %>% 
  ggplot(aes(order, freq)) +
    geom_area(aes(fill = reorder(hex, !!orderby))) +
  scale_fill_identity() +
  scale_color_identity() +
  equal_margins()
```

## (Optional) Smoothing

The landscape is too complex and you want to smooth out the wiggles.  This approach uses a loess smoother to average frequencies for each color.  Adjust the span closer to 1 for more smoothing, closer to zero for less:
```{r smoothing-param}
span <- 0.15 
n_points <- 1000 # number of points on x-axis
```

```{r smooth-freq}
order_grid <- seq(0, max(route_images$order), length.out = n_points)
route_smooth <- 
  route_images %>% 
  group_by(hex, L, U, V, H) %>% 
  nest() %>% 
  mutate(
   smooth_fun = map(data, ~ 
      loess(sqrt(freq) ~ order, data = ., span = span)),
   smooth = map(smooth_fun, 
     ~ tibble(
        order = order_grid, 
        freq = predict(., newdata = order_grid)^2)
      )
  ) %>% 
  unnest(smooth)  
```

There is no guarantee the areas add to a constant, so this often gives a wavy top and bottom:
```{r smooth-plot, fig.width = 10, fig.height = 3}
route_smooth  %>% 
  ggplot(aes(order, freq)) +
    geom_area(aes(fill = reorder(hex, !!orderby), color = hex)) +
  scale_fill_identity() +
  scale_color_identity() +
  equal_margins()
```

Scaling to have frequencies add to 1 restores the rectangular boundaries:
```{r smooth-scaled-plot, fig.width = 10, fig.height = 3}
route_smooth  %>% 
  group_by(order) %>% 
  mutate(freq = freq/(sum(freq))) %>% 
  ggplot(aes(order, freq)) +
    geom_area(aes(fill = reorder(hex, !!orderby), color = hex)) +
  scale_fill_identity() +
  scale_color_identity() +
  equal_margins() 
```

## Save image

`height` and `width` are in inches, `dpi = 300` is good for professional printing:
```{r save-plot}
ggsave(here("{{{ shortname }}}", "{{{ shortname }}}_route.jpeg"), 
  height = 6, width = 20, dpi = 300)
```

